From DSL to HPC Component-Based Runtime: A Multi-Stencil DSL Case Study

### 2016-01-20 ARTICLES/OTHER/LaMHA_261115/slides.tex
clean Helene 20/01/2016

%-------------------------------------------------------------------
\title[From DSL to HPC Component-Based Runtime: A Multi-Stencil DSL Case Study]{From DSL to HPC Component-Based Runtime: A Multi-Stencil DSL Case Study}
\author[Julien Bigot (CEA), Hélène Coullon (INRIA), Christian Perez (INRIA)]{Julien Bigot, \underline{Hélène Coullon}, Christian Perez}
\institute[INRIA]{INRIA team Avalon\\Maison de la simulation (CEA)}
\date{$16^{th}$ November 2015}
%-------------------------------------------------------------------

@@

% INTRODUCTION / MOTIVATION
%-------------------------------------------------------------
\begin{frame}
\frametitle{Motivation} % HPC DSL
\begin{block}{+ Domain Specific Languages}
\begin{itemize}
\item Easy language for the user
\item Implicit optimizations
\item Implicit parallelization
\end{itemize}
\end{block}
\begin{alertblock}{- Domain Specific Languages}
\begin{itemize}
\item Difficulties deported to the DSL designer
\begin{itemize}
\item Low level high performance programming
\item Maintainability and portability
\end{itemize}
\item As many DSLs as domains
\begin{itemize}
\item Productivity and code-reuse
\item Separation of concerns
\end{itemize}
\end{itemize}
\end{alertblock}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Motivation} % components
\begin{block}{Component models}
\begin{itemize}
\item Divide an application into several functionnal black boxes
\item Each component defines its connections with outer world
\item Each component are built independently from the others
\item Application = Assembly of components
\end{itemize}
\end{block}
\begin{block}{+ Component models}
\begin{itemize}
\item Maintainability through separation of concerns
\item Code-reuse and productivity
\item Dynamic assembly of components
\end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Motivation} % components
What if a DSL produces a component-based runtime ?

\begin{itemize}
\item Is it efficient ?
\item Is it feasible ?
\item Does it improve issues of DSLs ?
\begin{itemize}
\item maintainability
\item portability
\item productivity
\end{itemize}
\end{itemize}

@@

  \node[purp2,align=center] (edp) at ({360/\n * (2 - 1)}:\radius) {Partial Differential Equations\\\scriptsize $\frac{\partial u(x,y,t)}{\partial t} = \frac{\partial^2 u(x,y,t)}{\partial x^2} + \frac{\partial^2 u(x,y,t)}{\partial y^2}$};
  \node[progm,align=center,right of=edp,node distance=6cm] (cond) {\scriptsize + specific behavior for\\\scriptsize boundary conditions};
  \path[line,dashed] (edp) -- (cond);
  \node[purp,align=center] (maillage) at ({360/\n * (1 - 1)}:\radius) {Mesh\\Time iterations};
  \path[->] (maillage) edge[<-,thin,shorten <=2pt,shorten >=2pt, bend right=40]  node [fill=white,align=center] {\scriptsize Time and space\\\scriptsize discretization} (edp.east);
  \path[] (edp.west) edge[thin,shorten <=2pt,shorten >=2pt, bend right=90]  node [draw,circle,fill=white] (plus) {\scriptsize +} (maillage.west);
  \node[purp,align=center, below of=plus, node distance=3.5cm] (sche) {Explicit\\numerical schemes\\\textcolor{red}{stencils}};
  \node[progm,align=center,right of=sche,node distance=5.5cm] (expl) {\scriptsize Step by step approximation of the phenomena};
  \path[line,dashed] (sche) -- (expl);
  \path[line] (plus) -- node [fill=white,align=center] {\scriptsize Numerical mathods\\\scriptsize Finite difference/volume/elment} (sche);
\end{tikzpicture}
}
\end{center}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Time and Mesh}
\begin{block}{Time}
At each time iteration of the simulation are applied the \emph{computations kernels} of the application.
\end{block}
\begin{block}{Mesh}
\begin{itemize}
\item A Mesh is a connected undirected graph $\mathcal{M}=(V,E)$ without bridges
\item Mesh entities are a subset of $V \cup E$
\end{itemize}
\end{block}
\begin{center}
\resizebox{6cm}{!}{\includegraphics{./images/mesh.pdf}}
\end{center}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Data and Computation Kernels}
\begin{block}{Data}
A Data is a set of numerical values, each one attached to a given mesh entity
\end{block}
\begin{block}{Computation kernel}
\begin{itemize}
\item Set of data read for the computation
\begin{itemize}
\item Each one associated to a stencil shape
\end{itemize}
\item Data written by the computaion
\item A numerical expression
\item A computation domain
\begin{itemize}
\item Subset of mesh entities
\end{itemize}
\end{itemize}
\end{block}
\end{frame}
%----------------

@@

%----------------
\begin{frame}
\frametitle{Computation kernel: Examples}
\begin{center}
\resizebox{6cm}{!}{\includegraphics{./images/mesh.pdf}}
\hspace{10pt}
\resizebox{5cm}{!}{\includegraphics{./images/stencil2.pdf}}
\end{center}
Compute $A$ using $\{(A,local),(C,n2)\}$ on the computation domain called $dc2$ with the numerical expression
\tiny
\begin{equation*}
A(x,y)=A(x,y)+C(x1,y1)+C(x1+1,y1).
\end{equation*}
\end{frame}
%----------------

@@

%----------------
\begin{frame}
\frametitle{Multi-Stencil Language}
\begin{alertblock}{MSL is not}
\begin{itemize}
\item a new stencil optimizer/compiler
\item a new distributed data structure
\end{itemize}
\end{alertblock}
\begin{block}{MSL is}
\begin{itemize}
\item a high-level language for multi-stencil simulations
\item agnostic from the type of mesh used (data structure)
\item based on identifiers only
\end{itemize}
\end{block}
\begin{center}
\textit{MSL produces a ready-to-fill\\component-based parallel pattern of the simulation !}
\end{center}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Related Work}
\begin{block}{Complementary work}
\begin{itemize}
\item Distributed data structures: SkelGIS, Global Arrays
\item Stencil DSLs (on grids): Pochoir, PATUS
\item Stencil DSLs (on unstructured meshes): OP2, Liszt
\end{itemize}
\end{block}
\begin{alertblock}{Similar work}
\begin{itemize}
\item Pipeline of stencil computations for image processing: Halide
\item DSL to component-based runtime: ?
\end{itemize}
\end{alertblock}
\end{frame}
%----------------
%-------------------------------------------------------------

@@

%----------------
\begin{frame}
\frametitle{MSL to Component-based runtime}
\begin{block}{Ready-to-fill parallel pattern}
\begin{itemize}
\item Data parallelism
\begin{itemize}
\item External distributed data structure
\item Automatic detection of synchronizations
\end{itemize}
\item Task parallelism (mid-grain)
\begin{itemize}
\item Compile a static scheduling of computation kernels
\end{itemize}
\end{itemize}
\end{block}
The fine grain task parallelism is left to other languages: 
\begin{itemize}
\item OpenMP in the kernels
\item Kernels generated by stencil compilers (Pochoir, PATUS, Liszt etc.)
\end{itemize}
\end{frame}
%----------------

@@

%-------------------------------------------------------------
\section{Compiler}
% A slide example MS, 6 slides on graph transformations
%-------------------------------------------------------------
\subsection{Compiler}
%----------------
\begin{frame}[fragile]
\frametitle{Example}
\begin{columns}
\begin{column}{.52\textwidth}
\begin{lstlisting}[style=base]
@mesh:@ cart
@mesh entities:@ cell,edgex,edgey
@computation domains:@
  allcell in cell
  alledgex in edgex
  alledgey in edgey
  part1edgex in edgex
  part2edgex in edgex
@data:@
  a,cell
  b,cell
  c,edgex
  d,edgex
  e,edgey
\end{lstlisting}
\end{column}
\begin{column}{.48\textwidth}
\begin{lstlisting}[style=base]
  f,cell
  g,edgey
  h,edgex
  i,cell
  j,edgex
@time:@500
@computations:@
  b[allcell]=&c0&(a)
  c[alledgex]=&c1&(b[n1])
  d[alledgex]=&c2&(c)
  e[alledgey]=&c3&(c)
  f[allcell]=&c4&(d[n1])
  g[alledgey]=&c5&(e)
  h[alledgex]=&c6&(f)
  i[allcell]=&c7&(g,h)
  j[partedgex]=&c8&(i[n1])
\end{lstlisting}
\end{column}
\end{columns}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Data parallelism}
$\Gamma = [c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8]$\\
$\hookrightarrow [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$

\medskip
\begin{block}{Synchronization}
For $n$ the number of computations in $\Gamma$, and for $i,j$ such that $i<j<n$, a \textit{synchronization} is needed between $c_i$ and $c_j$, denoted $c_i \pprec c_j$, if $c_j$ read a data written by $c_i$ with a stencil shape and not locally.
\end{block}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Mid-grain task parallelism}

Build the read/write dependency graph.

@@

\begin{center}
\textbf{Data parallelism + Task parallelism}
\end{center}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Which scheduling ?}
From the dependency graph we can:
\begin{itemize}
\item Use a dynamic scheduler
\item Compute a static scheduling
\end{itemize}

In this work,
\begin{itemize}
\item We build a static scheduling of the multi-stencil application
\item This static scheduling can be dumped to a component-based runtime
\end{itemize}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Series-Parallel Tree}
\begin{columns}
\begin{column}{.48\textwidth}
\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0)    at (0,0) {$\mathcal{S}$};

@@

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (c2)
         edge node {} (star4)
         edge node {} (c4)
         edge node {} (c4)
         edge node {} (c6)
    (s2) edge node {} (c3)
         edge node {} (c5);
  \end{tikzpicture}
  }
\end{center}
\end{column}
\begin{column}{.52\textwidth}
\begin{block}{Specific components}
\begin{itemize}
\item $SEQ$ to directly replace $\mathcal{S}$ nodes
\item $PAR$ to directly replace $\mathcal{P}$ nodes
\item $SYNC$ for synchronizations
\item $K$ for computation kernels
\end{itemize}
\end{block}
\end{column}
\end{columns}
\begin{center}
\textbf{Direct dump to the component assembly}
\end{center}
\end{frame}
%----------------

@@

%----------------
\begin{frame}
\frametitle{Back to Data Parallelism}
\begin{center}
\textbf{Back to Data Parallelism only ?}\\
\end{center}
\textbf{Loop Fusion}: Same computation domain under a sequence or a parallel node

\medskip
\begin{center}
\resizebox{0.5\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0)    at (0,0) {$\mathcal{S}$};

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (c2)
         edge node {} (star4)
         edge node {} (c4)
         edge node {} (c4)
         edge node {} (c6)
    (s2) edge node {} (c3);
  \end{tikzpicture}
  }
\end{center}

\small
\textbf{Canonical form}: $[c_0,sync_1,c_1,c_2,sync_4,c_4,c_6,\{c_3,c_5\},c_7,sync_8,c_8]$
\end{frame}
%----------------

@@

%-------------------------------------------------------------
\section{Evaluation}
% Present Shallow-water case, present results
%-------------------------------------------------------------
\subsection{Evaluation}
\begin{frame}
\frametitle{Implementation and evaluation}
\textbf{Implementation of MSL}: Python, SkelGIS and $L^2C$

@@

\medskip
\textbf{Evaluation of the data parallelism}
\begin{itemize}
\item Full SkelGIS implementation (DDS + specific interfaces to hide communications)
\item MSL implementation which uses the SkelGIS DDS
\item Cluster Edel of Grid'5000: 2 8-cores Intel Xeon E5520, 24GB RAM / nodes
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Evaluations}
\begin{columns}
\begin{column}{.5\textwidth}
\begin{center}
\small Mesh $10k \times 10k$, $500$ iterations\\
\resizebox{\textwidth}{!}{\includegraphics{./images/logtimes_10K_500.pdf}}
\end{center}
\end{column}
\begin{column}{.5\textwidth}
\begin{center}
\small Mesh $10k \times 10k$, $2K$ iterations\\
\resizebox{\textwidth}{!}{\includegraphics{./images/logtimes_10K_2K.pdf}}
\end{center}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\section{Conclusion and perspectives}
%-------------------------------------------------------------
\subsection{Conclusion and perspectives}
\begin{frame}
\frametitle{Conclusion}
\begin{itemize}
\item A DSL for Multi-Stencil applications (MSL)
\item The compilation of MSL to get a static scheduling
\item The dump to a component-based runtime
\item No overhead introduced by the solution
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Perspectives}
\begin{itemize}
\item A comparison with an improved version of SkelGIS
\item More evaluations on TGCC Curie (CEA) up to $32k$ cores
\item An evaluation of the Data+Task parallelism (mixing the static scheduling and OpenMP)
\item A new dump with a component responsible for the dynamic scheduling of the application (using OpenMP 4.x)
\item A new dump to CPU+GPGPUs (using existing stencil compilers Pochoir, PATUS etc.)
\item An alternative to the SkelGIS DDS using Global Arrays (PGAS)
\end{itemize}
\end{frame}
%-------------------------------------------------------------

\begin{withoutheadline}
\begin{frame}{}
\begin{center}
\huge Thank You !
\end{center}
\end{frame}
\end{withoutheadline}

>>>>

%-------------------------------------------------------------------
\title[From DSL to HPC Component-Based Runtime: A Multi-Stencil DSL Case Study]{From DSL to HPC Component-Based Runtime: A Multi-Stencil DSL Case Study}
\author[Julien Bigot (CEA), Hélène Coullon (INRIA), Christian Perez (INRIA)]{Julien Bigot, \underline{Hélène Coullon}, Christian Perez}
\institute[INRIA]{INRIA team Avalon\\Maison de la simulation (CEA)}
\date{WOLFHPC 2015 - $16^{th}$ November 2015}
%-------------------------------------------------------------------

@@

% INTRODUCTION / MOTIVATION
%-------------------------------------------------------------
\begin{frame}
\frametitle{Motivation} % HPC DSL
\begin{block}{+ Domain Specific Languages}
\begin{itemize}
\item Separation of concerns (domain/implementation)
\item Easy language for the user
\item Implicit optimizations
\item Implicit parallelization
\end{itemize}
\end{block}
\begin{alertblock}{- Domain Specific Languages}
\begin{itemize}
\item Difficulties deported to the DSL designer
\begin{itemize}
\item Low level high performance programming
\item Maintainability and portability
\end{itemize}
\item As many DSLs as domains
\begin{itemize}
\item DSL composition ?
\end{itemize}
\end{itemize}
\end{alertblock}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Motivation} % components
\begin{block}{Component models}
\begin{itemize}
\item Divide an application into several independent black boxes
\item Each component defines its interactions with outer world
\item Application = Assembly of components
\end{itemize}
\end{block}
\begin{block}{+ Component models}
\begin{itemize}
\item Maintainability through separation of concerns
\item Code-reuse and productivity
\item Dynamic assembly of components
\end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Motivation} % components
What if a DSL produces a component-based runtime ?

\begin{itemize}
\item Is it feasible ?
\item Is it efficient ?
\item \textcolor{gray}{Does it improve issues of DSLs ?}
\begin{itemize}
\item \textcolor{gray}{maintainability}
\item \textcolor{gray}{portability}
\item \textcolor{gray}{productivity}
\end{itemize}
\end{itemize}

@@

  \node[purp2,align=center] (edp) at ({360/\n * (2 - 1)}:\radius) {Partial Differential Equations\\\scriptsize $\frac{\partial u(x,y,t)}{\partial t} = \frac{\partial^2 u(x,y,t)}{\partial x^2} + \frac{\partial^2 u(x,y,t)}{\partial y^2}$};
  \node[progm,align=center,right of=edp,node distance=6cm] (cond) {\scriptsize + specific behavior for\\\scriptsize boundary conditions};
  \path[line,dashed] (edp) -- (cond);
  \node[purp,align=center] (maillage) at ({360/\n * (1 - 1)}:\radius) {Mesh\\Time iterations};
  \path[->] (maillage) edge[<-,thin,shorten <=2pt,shorten >=2pt, bend right=40]  node [fill=white,align=center] {\scriptsize Time and space\\\scriptsize discretization} (edp.east);
  \path[] (edp.west) edge[thin,shorten <=2pt,shorten >=2pt, bend right=90]  node [draw,circle,fill=white] (plus) {\scriptsize +} (maillage.west);
  \node[purp,align=center, below of=plus, node distance=3.5cm] (sche) {Explicit\\numerical schemes\\\textcolor{red}{stencils}};
  \node[progm,align=center,right of=sche,node distance=5.5cm] (expl) {\scriptsize Step by step approximation of the phenomena};
  \path[line,dashed] (sche) -- (expl);
  \path[line] (plus) -- node [fill=white,align=center] {\scriptsize Numerical methods\\\scriptsize Finite difference/volume/elment} (sche);
\end{tikzpicture}
}
\end{center}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Time and Mesh}
\begin{block}{Time}
At each time iteration of the simulation are applied the \emph{computation kernels} of the application.
\end{block}
\begin{block}{Mesh}
\begin{itemize}
\item A Mesh is a connected undirected graph $\mathcal{M}=(V,E)$ without bridges
\item Mesh entities are a subset of $V \cup E$
\end{itemize}
\end{block}
\begin{center}
\resizebox{6cm}{!}{\includegraphics{./images/mesh.pdf}}
\end{center}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Data and Computation Kernels}
\begin{block}{Data}
Data is a set of numerical values, each one attached to a given mesh entity
\end{block}
\begin{block}{Computation kernel}
\begin{itemize}
\item Set of data read for the computation
\begin{itemize}
\item Each one associated to a stencil shape
\end{itemize}
\item Data written by the computation
\item A numerical expression
\item A computation domain
\begin{itemize}
\item Subset of mesh entities
\end{itemize}
\end{itemize}
\end{block}
\end{frame}
%----------------

@@

%----------------
% \begin{frame}
% \frametitle{Computation kernel: Example}
% \begin{center}
% \resizebox{6cm}{!}{\includegraphics{./images/mesh.pdf}}
% \hspace{10pt}
% \resizebox{5cm}{!}{\includegraphics{./images/stencil2.pdf}}
% \end{center}
% \end{frame}
%----------------

@@

%----------------
\begin{frame}[fragile]
\frametitle{Example}
\begin{columns}
\begin{column}{.52\textwidth}
\resizebox{\textwidth}{!}{\includegraphics{./images/mesh.pdf}}
\hspace{10pt}
\resizebox{\textwidth}{!}{\includegraphics{./images/stencil2.pdf}}
\end{column}
\begin{column}{.48\textwidth}
$\mathcal{MSP}(T,\mathcal{M},\mathcal{E},\mathcal{D},\Delta,\Gamma)$\\

\begin{lstlisting}[style=base]
@mesh:@ cart
@mesh entities:@ cell@,@edgex
@computation domains:@
  allcell @in@ cell
  alledgex @in@ edgex
@data:@
  A@,@cell
  C@,@edgex
@time:@500
@computations:@
  A@[@allcell@]@=&comp&@(@C@[@n1@]@@)@
\end{lstlisting}
\end{column}
\end{columns}
\end{frame}

% %----------------
% \begin{frame}[fragile]
% \frametitle{MSL Example}
% \begin{columns}
% \begin{column}{.52\textwidth}
% \begin{lstlisting}[style=base]
% @mesh:@ cart
% @mesh entities:@ cell,edgex,edgey
% @computation domains:@
%   allcell in cell
%   alledgex in edgex
%   alledgey in edgey
%   part1edgex in edgex
%   part2edgex in edgex
% @data:@
%   a,cell
%   b,cell
%   c,edgex
%   d,edgex
%   e,edgey
% \end{lstlisting}
% \end{column}
% \begin{column}{.48\textwidth}
% \begin{lstlisting}[style=base]
%   f,cell
%   g,edgey
%   h,edgex
%   i,cell
%   j,edgex
% @time:@500
% @computations:@
%   b[allcell]=&c0&(a)
%   c[alledgex]=&c1&(b[n1])
%   d[alledgex]=&c2&(c)
%   e[alledgey]=&c3&(c)
%   f[allcell]=&c4&(d[n1])
%   g[alledgey]=&c5&(e)
%   h[alledgex]=&c6&(f)
%   i[allcell]=&c7&(g,h)
%   j[partedgex]=&c8&(i[n1])
% \end{lstlisting}
% \end{column}
% \end{columns}
% \end{frame}
% %----------------

%----------------
\begin{frame}
\frametitle{Multi-Stencil Language}
\begin{alertblock}{MSL is not}
\begin{itemize}
\item a new stencil optimizer/compiler
\item a new distributed data structure
\end{itemize}
\end{alertblock}
\begin{block}{MSL is}
\begin{itemize}
\item a high-level language for multi-stencil simulations
\item agnostic from the type of mesh used (data structure)
\item based on identifiers only
\end{itemize}
\end{block}
\begin{center}
\textit{MSL produces a "ready-to-fill"\\component-based parallel scheduling of the simulation}
\end{center}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Related Work}
\begin{block}{Complementary work}
\begin{itemize}
\item Distributed data structures: SkelGIS, Global Arrays
\item Stencil DSLs (on grids): Pochoir, PATUS
\item Stencil DSLs (on unstructured meshes): OP2, Liszt
\end{itemize}
\end{block}
\begin{alertblock}{Similar work}
\begin{itemize}
\item The SIPSim model (DDS,Data,applicators and iterators)
\begin{itemize}
\item Abstraction of a distributed data structure
\end{itemize}
\item Pipeline of stencil computations for image processing: Halide
\begin{itemize}
\item On grids (image), different abstraction level
\end{itemize}
\item DSL to component-based runtime: ?
\end{itemize}
\end{alertblock}
\end{frame}
%----------------
%-------------------------------------------------------------

@@

%----------------
\begin{frame}
\frametitle{MSL to Component-based runtime}
\begin{block}{Ready-to-fill parallel scheduling: mid-grain parallelism}
\begin{itemize}
\item Data parallelism
\begin{itemize}
\item External distributed data structure
\item Automatic detection of synchronizations
\end{itemize}
\item Task parallelism
\begin{itemize}
\item Compile a static scheduling of computation kernels
\end{itemize}
\end{itemize}
\end{block}
The fine grain parallelism is left to other languages: 
\begin{itemize}
\item OpenMP in the kernels
\item Kernels generated by stencil compilers for CPU or GPU (Pochoir, Liszt etc.)
\end{itemize}
\end{frame}
%----------------

@@

%-------------------------------------------------------------
\section{Compiler}
% A slide example MS, 6 slides on graph transformations
%-------------------------------------------------------------
\subsection{Compiler}
%----------------
\begin{frame}[fragile]
\frametitle{Example}
\begin{columns}
\begin{column}{.52\textwidth}
\begin{lstlisting}[style=base]
@mesh:@ cart
@mesh entities:@ cell,edgex,edgey
@computation domains:@
  allcell @in@ cell
  alledgex @in@ edgex
  alledgey @in@ edgey
  part1edgex @in@ edgex
  part2edgex @in@ edgex
@data:@
  a,cell
  b,cell
  c,edgex
  d,edgex
  e,edgey
\end{lstlisting}
\end{column}
\begin{column}{.48\textwidth}
\begin{lstlisting}[style=base]
  f,cell
  g,edgey
  h,edgex
  i,cell
  j,edgex
@time:@500
@computations:@
  b[allcell]=&c0&(a)
  c[alledgex]=&c1&(b[n1])
  d[alledgex]=&c2&(c)
  e[alledgey]=&c3&(c)
  f[allcell]=&c4&(d[n1])
  g[alledgey]=&c5&(e)
  h[alledgex]=&c6&(f)
  i[allcell]=&c7&(g,h)
  j[partedgex]=&c8&(i[n1])
\end{lstlisting}
\end{column}
\end{columns}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Data parallelism}
\begin{enumerate}
\item Assembly of components duplicated on each resource
\item External Distributed Data Structure to split data among resources
\item Detect when synchronizations are needed
\end{enumerate}

\medskip
\begin{block}{Synchronization}
When a computation read a data, usign a stencil shape, that has been written by a previous computation.
\end{block}

$\Gamma = [c_0,c_1,c_2,c_3,c_4,c_5,c_6,c_7,c_8]$\\
$\hookrightarrow [c_0,sync_1,c_1,c_2,c_3,sync_4,c_4,c_5,c_6,c_7,sync_8,c_8]$
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Data and task parallelism}

\begin{block}{Dependency graph}
\begin{enumerate}
\item Each node is a computation or a synchronization
\item Each edge is a dependency: a computation read a data that has been written before.
\end{enumerate}
\end{block}

@@

\begin{center}
\textcolor{gray}{Dynamic} or static scheduling ?
\end{center}
\end{frame}
%----------------

%----------------
% \begin{frame}
% \frametitle{Which scheduling ?}
% From the dependency graph we can:
% \begin{itemize}
% \item Use a dynamic scheduler
% \item Compute a static scheduling
% \end{itemize}

% In this work,
% \begin{itemize}
% \item We build a static scheduling of the multi-stencil application
% \item This static scheduling can be dumped to a component-based runtime
% \end{itemize}
% \end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Series-Parallel Tree}
\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
\normalsize
\begin{columns}
\begin{column}{.48\textwidth}
\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0)    at (0,0) {$\mathcal{S}$};

@@

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (c2)
         edge node {} (star4)
         edge node {} (c4)
         edge node {} (c4)
         edge node {} (c6)
    (s2) edge node {} (c3)
         edge node {} (c5);
  \end{tikzpicture}
  }
\end{center}
\end{column}
\begin{column}{.52\textwidth}

\end{column}
\end{columns}
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Series-Parallel Tree}
\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
\normalsize
\begin{columns}
\begin{column}{.48\textwidth}
\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0)    at (0,0) {$\mathcal{S}$};

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (c2)
         edge node {} (star4)
         edge node {} (c4)
         edge node {} (c4)
         edge node {} (c6)
    (s2) edge node {} (c3);
  \end{tikzpicture}
  }
\end{center}
\end{column}
\begin{column}{.52\textwidth}
\end{column}
\end{columns}
Loop fusion optimization possible
\end{frame}
%----------------

%----------------
\begin{frame}
\frametitle{Series-Parallel Tree}
\small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
\normalsize
\begin{columns}
\begin{column}{.48\textwidth}
\begin{center}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (s0)    at (0,0) {$\mathcal{S}$};

  \path[->]
    (s0) edge node {} (c0)
         edge node {} (star1)
         edge node {} (c1)
         edge node {} (p0)
         edge node {} (c7)
         edge node {} (star8)
         edge node {} (c8)
    (p0) edge node {} (s1)
         edge node {} (s2)
    (s1) edge node {} (c2)
         edge node {} (star4)
         edge node {} (c4)
         edge node {} (c4)
         edge node {} (c6)
    (s2) edge node {} (c3);
  \end{tikzpicture}
  }
\end{center}
\end{column}
\begin{column}{.52\textwidth}
\begin{block}{Specific components}
\begin{itemize}
\item $SEQ$ to directly replace $\mathcal{S}$ nodes
\item $PAR$ to directly replace $\mathcal{P}$ nodes
\item $SYNC$ for synchronizations
\item $K$ for computation kernels
\end{itemize}
\end{block}
\end{column}
\end{columns}
Loop fusion optimization possible
\end{frame}
%----------------

@@

%----------------
% \begin{frame}
% \frametitle{Data Parallelism optimization}
% \begin{block}{Loop Fusion}
% Same computation domain under a sequence or a parallel node
% \end{block}

% \medskip
% \begin{center}
% \resizebox{0.5\textwidth}{!}{%
% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
%    \node[] (s0)    at (0,0) {$\mathcal{S}$};

%    \node[] (c0)    at (-3,1) {$c_0$};
%    \node[] (star1) at (-2,1) {$sync_1$};
%    \node[] (c1)    at (-1,1) {$c_1$};
%    \node[] (p0)    at ( 0,1) {$\mathcal{P}$};
%    \node[] (c7)    at ( 1,1) {$c_7$};
%    \node[] (star8) at ( 2,1) {$sync_8$};
%    \node[] (c8)    at ( 3,1) {$c_8$};

%    \node[] (s1)    at (-1,2) {$\mathcal{S}$};
%    \node[] (s2)    at (1,2) {$\mathcal{S}$};

%    \node[] (c2)    at (-3,3) {$c_2$};
%    \node[] (star4) at (-2,3) {$sync_4$};
%    \node[] (c4)    at (-1,3) {$c_4$};
%    \node[] (c6)    at ( 0,3) {$c_6$};
%    \node[] (c3)    at ( 1,3) {\textcolor{red}{$c_3;c_5$}};
%    %\node[] (c5)    at ( 2,3) {$c_5$};

%   \path[->]
%     (s0) edge node {} (c0)
%          edge node {} (star1)
%          edge node {} (c1)
%          edge node {} (p0)
%          edge node {} (c7)
%          edge node {} (star8)
%          edge node {} (c8)
%     (p0) edge node {} (s1)
%          edge node {} (s2)
%     (s1) edge node {} (c2)
%          edge node {} (star4)
%          edge node {} (c4)
%          edge node {} (c4)
%          edge node {} (c6)
%     (s2) edge node {} (c3);
%   \end{tikzpicture}
%   }
% \end{center}

% \small
% %\textbf{Canonical form}: $[c_0,sync_1,c_1,c_2,sync_4,c_4,c_6,\{c_3,c_5\},c_7,sync_8,c_8]$
% \end{frame}
%----------------

@@

%-------------------------------------------------------------
\section{Evaluation}
% Present Shallow-water case, present results
%-------------------------------------------------------------
\subsection{Evaluation}

%----------------
\begin{frame}
\frametitle{Resume}
The MSL compiler can produce:
\begin{itemize}
\item A data parallel pattern of the multi-stencil application
\item \textcolor{gray}{An hybrid (data + task) pattern of the multi-stencil application}
\end{itemize}
\end{frame}
%----------------

\begin{frame}
\frametitle{Implementation and evaluation}
\textbf{Implementation of MSL}: Python, SkelGIS and $L^2C$

@@

\medskip
\textbf{Evaluation of the data parallelism}
\begin{itemize}
\item Full SkelGIS implementation (DDS + specific interfaces to hide communications)
\item MSL implementation which uses the SkelGIS DDS
%\item Cluster Edel of Grid'5000: 2 8-cores Intel Xeon E5520, 24GB RAM / nodes, Infiniband
\item Thin Nodes TGCC Curie: two 8-cores Intel Sandy Bridge 2.7GHz, 64GB RAM, Infiniband
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Evaluations}
Mesh size: $10k \times 10k$
Number of iterations: $500$
\begin{columns}
\begin{column}{.5\textwidth}
\begin{center}
\resizebox{\textwidth}{!}{\includegraphics{./images/logtimes_NF.pdf}}
\end{center}
\end{column}
\begin{column}{.5\textwidth}
\begin{center}
\resizebox{\textwidth}{!}{\includegraphics{./images/itpersec_NF.pdf}}
\end{center}
\end{column}
\end{columns}
\end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\section{Conclusion and perspectives}
%-------------------------------------------------------------
\subsection{Conclusion and perspectives}
\begin{frame}
\frametitle{Conclusion}
\begin{block}{Conclusion}
\begin{itemize}
\item A DSL for Multi-Stencil applications (MSL)
\item The compilation of MSL to get a parallel scheduling pattern of the simulation
\begin{itemize}
\item Data parallelism
\item Task parallelism
\end{itemize}
\item The dump to a component-based runtime
\item Data parallelism evaluation: no overhead introduced 
\end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------
